### 2024-11-08 final-project.Rmd
table 2 and diagnostics

---
title: "final-project"
author: "Murphy John"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(stringr)
library(psych)
library(gmodels)
library(survival)
library(survminer)
```

```{r}
# load data
seer1 <- read.csv("data/caselist_2024.11.csv", header = TRUE)
```

```{r}
# column rename
seer_name <- seer1 %>%
  dplyr::rename(
  age = Age.recode.with..1.year.olds,
  sex = Sex,
  year_diagnosis = Year.of.diagnosis,
  race = Race.ethnicity,
  marriage_status = Marital.status.at.diagnosis,
  tnm2015 = Derived.AJCC.T..7th.ed..2010.2015.,
  tnm2017 = Derived.SEER.Combined.T..2016.2017.,
  tnm2018 = Derived.EOD.2018.T..2018..,
  subtype = Breast.Subtype..2010..,
  grade2017 = Grade.Recode..thru.2017.,
  grade2018 = Derived.Summary.Grade.2018..2018..,
  n2015 = Derived.AJCC.N..7th.ed..2010.2015.,
  n2017 = Derived.SEER.Combined.N..2016.2017.,
  n2018 = Derived.EOD.2018.N..2018..,
  m2015 = Derived.AJCC.M..7th.ed..2010.2015.,
  m2017 = Derived.SEER.Combined.M..2016.2017.,
  m2018 = Derived.EOD.2018.M..2018..,
  radiation = Radiation.recode,
  chemotherapy = Chemotherapy.recode..yes..no.unk.,
  surgery = RX.Summ..Surg.Prim.Site..1998..,
  survival_months = Survival.months,
  survival_status = Vital.status.recode..study.cutoff.used.,
  survival_cause_cancer = SEER.cause.specific.death.classification,
  survival_cause_other = SEER.other.cause.of.death.classification,
  idc_type = Histologic.Type.ICD.O.3
)

# filter
seer_filter <- seer_name %>%
  filter(
    age != "00 years",
    age != "15-19 years",
    sex == "Female",
    survival_months >= 1,
    survival_cause_cancer != "Dead (missing/unknown COD)",
    survival_cause_other != "Dead (attributable to causes other than this cancer dx)",
    grade2017 != "B-cell; pre-B; B-precursor",
    grade2017 != "T-cell",
    idc_type %in% seq(8500, 8549)
  )

# clean up the variables
seer_clean <- seer_filter %>%
  mutate(
    # assign race as White, Black, or Other
    race = case_when(
      race == "White" ~ "White",
      race == "Black" ~ "Black",
      TRUE ~ "Other"
    ),
    # assign marriage_status as Yes or No
    marriage_status = case_when(
      marriage_status == "Married (including common law)" ~ "Yes",
      TRUE ~ "No"
    ),

    # assign blank tnm to NA
    tnm2015 = case_when(
      tnm2015 == "Blank(s)" ~ NA_character_,
      TRUE ~ tnm2015
    ),
    tnm2017 = case_when(
      tnm2017 == "Blank(s)" ~ NA_character_,
      TRUE ~ tnm2017
    ),
    tnm2018 = case_when(
      tnm2018 == "Blank(s)" ~ NA_character_,
      TRUE ~ tnm2018
    ),

    # merge tnm to ensure cross-validation
    tnm = coalesce(tnm2015, tnm2017, tnm2018),

    # assign breast cancer subtypes to their more commonly used names
    subtype = case_when(
      subtype == "HR-/HER2-" ~ "Triple Negative",
      subtype == "HR-/HER2+" ~ "HER2 Positive",
      subtype == "HR+/HER2-" ~ "Luminal A",
      subtype == "HR+/HER2+" ~ "Luminal B",
      TRUE ~ NA_character_
    ),
    # assign common grades between 2017 and 2018
    # "1" is well-differentiated, "2" is moderately-differentiated
    # "3" is poorly- differentiated, "4" is undifferentiated
    # see documentation for additional details on assignments
    grade2017 = case_when(
      grade2017 == "Well differentiated; Grade I" ~ "1",
      grade2017 == "Moderately differentiated; Grade II" ~ "2",
      grade2017 == "Poorly differentiated; Grade III" ~ "3",
      grade2017 == "Undifferentiated; anaplastic; Grade IV" ~ "4",
      grade2017 == "H" ~ "3",
      grade2017 == "Unknown" ~ NA_character_,
      TRUE ~ grade2017
    ),
    grade2018 = case_when(
      grade2018 == "A" ~ "1",
      grade2018 == "B" ~ "2",
      grade2018 == "C" ~ "3",
      grade2018 == "D" ~ "4",
      grade2018 == "H" ~ "3",
      grade2018 == "9" ~ NA_character_,
      grade2018 == "Blank(s)" ~ NA_character_,
      TRUE ~ grade2018
    ),
    # then merge the 2017 and 2018 grade columns
    grade = coalesce(grade2017, grade2018),

    # assign N stage type for easier analysis
    n2015 = case_when(
      str_detect(n2015, "N0") ~ "N0",
      str_detect(n2015, "N1|N1N0S") ~ "N1",
      str_detect(n2015, "N2|N2N0S") ~ "N2",
      str_detect(n2015, "N3|N3N0S") ~ "N3",
      str_detect(n2015, "NX") ~ NA_character_,
      TRUE ~ NA_character_
    ),
    n2017 = case_when(
      str_detect(n2017, "c0|p0") ~ "N0",
      str_detect(n2017, "c1|p1") ~ "N1",
      str_detect(n2017, "c2|p2") ~ "N2",
      str_detect(n2017, "c3|p3") ~ "N3",
      str_detect(n2017, "cX") ~ NA_character_,
      TRUE ~ NA_character_
    ),
    n2018 = case_when(
      str_detect(n2018, "N0") ~ "N0",
      str_detect(n2018, "N1|N1N0S") ~ "N1",
      str_detect(n2018, "N2|N2N0S") ~ "N2",
      str_detect(n2018, "N3|N3N0S") ~ "N3",
      str_detect(n2018, "NX") ~ NA_character_,
      TRUE ~ NA_character_
    ),
    # then merge the 2015, 2017, and 2018 N stage columns
    n_stage = coalesce(n2015, n2017, n2018),

     # assign M stage type for easier analysis
    m2015 = case_when(
      str_detect(m2015, "M0") ~ "M0",
      str_detect(m2015, "M1|M1M0S") ~ "M1",
      TRUE ~ NA_character_
    ),
    m2017 = case_when(
      str_detect(m2017, "c0") ~ "M0",
      str_detect(m2017, "c1|p1") ~ "M1",
      TRUE ~ NA_character_
    ),
    m2018 = case_when(
      str_detect(m2018, "M0") ~ "M0",
      str_detect(m2018, "M1|M1M0S") ~ "M1",
      TRUE ~ NA_character_
    ),

    # then merge the 2015, 2017, and 2018 M stage columns
    m_stage = coalesce(m2015, m2017, m2018),

    # simplify the radiation variable
    radiation = case_when(
      str_detect(radiation, "None|Refused|unknown") ~ "No/Unknown",
      TRUE ~ "Yes"
    ),
    surgery = case_when(
      surgery == "0" ~ "No/Unknown",
      surgery == "99" ~ "No/Unknown",
      TRUE ~ "Yes"
    )
    )

# there are several variables that need to be categorized for analysis and/filtered
seer_cat <- seer_clean %>%
   filter(
    !is.na(subtype),
    #!is.na(grade),
    #n_stage != "NX"
  ) %>%
  mutate(agecat = case_when(
    age %in% c("20-24 years", "25-29 years", "30-34 years", "35-39 years", "40-44 years", "45-49 years") ~ 0,
    age %in% c("50-54 years", "55-59 years", "60-64 years", "65-69 years") ~ 1,
    age %in% c("70-74 years", "75-79 years", "80-84 years", "85+ years") ~ 2,
    TRUE ~ NA_real_  # Keeps other values as NA if they don’t match any of the above groups
  )) %>%
  mutate(gradecat = case_when(
    grade %in% c(1, 2) ~ "I/II",
    grade %in% c(3, 4) ~ "III/IV",
    TRUE ~ NA_character_
    )) %>%
  # the survival_status column is what we use for event/censoring
  # recode Dead as 1 and Alive (censored) as 0
  mutate(status_cat = case_when(
    survival_status == "Dead" ~ 1,
    survival_status == "Alive" ~ 0
  ))

# further filter and select only variables used in analysis
seer <- seer_cat %>%
  select(
    age = agecat,
    sex,
    year_diagnosis,
    race,
    marriage_status,
    subtype,
    n_stage,
    m_stage,
    grade = gradecat,
    radiation,
    chemotherapy,
    surgery,
    survival_months,
    survival_status = status_cat,

  )

attach(seer)
```

```{r}
# Table 1: Clinicopathological Characteristics by Breast Cancer Subtype
# (unadjusted associations between covariates and exposure)

## age
CrossTable(age, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## race
CrossTable(race, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## marriage status
CrossTable(marriage_status, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## N stage
CrossTable(n_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## M stage
CrossTable(m_stage, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## grade
CrossTable(grade, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## radiation
CrossTable(radiation, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## chemotherapy
CrossTable(chemotherapy, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)

## surgery
CrossTable(surgery, subtype, prop.r = FALSE, prop.c = TRUE, prop.t = FALSE, prop.chisq = FALSE, chisq = TRUE)
```

```{r}
# Table 2: Clinicopathological Characteristics by survival
# (Unadjusted associations between covariates and survival)

## age
summary(survfit(Surv(time = survival_months, event = survival_status) ~ age), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ age)

## race
summary(survfit(Surv(time = survival_months, event = survival_status) ~ race), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ race)

## marriage status
summary(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ marriage_status)

## N stage
summary(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ n_stage)

## M stage
summary(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ m_stage)

## grade
summary(survfit(Surv(time = survival_months, event = survival_status) ~ grade), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ grade)

## radiation
summary(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ radiation)

## chemotherapy
summary(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ chemotherapy)

## surgery
summary(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), times = c(12, 60, 120))
survdiff(Surv(time = survival_months, event = survival_status) ~ surgery)
```

```{r}
# check the proportional hazards assumption for each covariate

## age
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ age), data = seer, fun="cloglog")

## race
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ race), data = seer, fun="cloglog")

## marriage status
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ marriage_status), data = seer, fun="cloglog")

## N stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ n_stage), data = seer, fun="cloglog")

## M stage
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ m_stage), data = seer, fun="cloglog")

## grade
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ grade), data = seer, fun="cloglog")

## radiation
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ radiation), data = seer, fun="cloglog")

## chemotherapy
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ chemotherapy), data = seer, fun="cloglog")

## surgery
ggsurvplot(survfit(Surv(time = survival_months, event = survival_status) ~ surgery), data = seer, fun="cloglog")
```