### 2024-11-25 products/manuscript/manuscript.qmd
manuscript start

---
title: "epid 8010 manuscript"
author: " "
format: pdf
bibliography:
  - "refs.bib"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	error = FALSE,
	knitr.graphics.error = FALSE
)
library(yaml, include.only = NULL)
library(knitr, include.only = NULL)
library(here)

# Loading data to present in text
seer <- readr::read_rds(here::here("data", "processed", "seer.rds"))
```

**Authors**

-   Murphy John$^{1}$

**Author affiliation**

1.  Department of Epidemiology and Biostatistics, University of Georgia, Athens, GA, USA.

\pagebreak

# Abstract

**Background:**

**Methods:**

**Results:**

**Conclusions:**

**Keywords:**

\pagebreak

# Introduction

Inflammatory breast cancer (IBC) is a rare but aggressive subtype of breast cancer, accounting for 2% to 6% of all breast cancer cases in the United States [@devi2019perspectives; @hance2005trends]. IBC presents differently than other types of breast cancer. It often lacks a breast lump and may not appear on a mammogram, making accurate diagnosis challenging and potentially delaying treatment. IBC causes symptoms of breast inflammation, such as swelling and redness, due to cancer cells blocking lymphatic vessels in the skin, giving the breast an "inflamed" appearance [@menta2018inflammatory].

Clinically, IBC is categorized as T4d in the TNM classification for its aggressive biological behavior. In spite of multimodal therapy, IBC has a local recurrence rate of up to 50% and survival rates of only 35% to 40%, significantly lower than those of other types of breast cancer [@robertson2010inflammatory]. Despite its rarity, IBC disproportionately accounts for up to 10% of all breast cancer-related deaths [@devi2019perspectives; @hance2005trends].

Because IBC is rare, the biological characteristics of IBC are not commonly reported and the molecular alterations that result in poor prognosis are not well understood. However, hormone receptors (HR) and human epidermal growth factor receptor-2 (HER2), which define the molecular status of IBC through immunohistochemistry (IHC), are fundamental markers used to demonstrate molecular features, predict prognosis, and optimize therapeutic regimens [@li2017outcomes]. IBC tumors are characterized by combinations of hormone receptors and oncogenes, allowing the majority to be characterized as HR+/HER2- (Luminal A), HR+/HER+ (Luminal B), HR-/HER2+ (HER2 Positive), and Triple Negative by IHC.[@TCR28003]. Studies have shown that these four molecular subtypes reveal the prognostic discrepancies in both common breast cancers and IBC [@robertson2010inflammatory; @le2015future; @tolaney2015adjuvant; @iwamoto2011different]. A previous study examined the association between molecular subtypes and survival outcomes in IBC patients from 2010 to 2013 [@li2017outcomes]. However, no updated population-based studies have investigated this relationship.

Therefore, using the most recent Surveillance, Epidemiology, and End Results (SEER) data, this study aims to assess the effect of IBC molecular subtype on breast cancer-specific mortality in adult females with IBC.

# Methods

The National Cancer Institute's Surveillance, Epidemiology, and End Results (SEER) database is an authoritative source of information on cancer incidence and survival in the United States. SEER currently collects and publishes cancer incidence and survival data from population-based cancer registries covering approximately 48% of the U.S. population. This study utilized SEER incidence data from 2010 to 2021.

## Study Population

The study population focuses on adult females diagnosed with IBC. Data was extracted from the SEER 17 Registries (2000-2021) using the SEER\*Stat software (version 8.4.4; Surveillance Research Program, NCI, Bethesda, MD). The specific criteria used to identify patients with IBC were as follows: (1) Site and Morphology (TNM 7/CS v0204 Schema recode) was limited to "Breast"; (2) Derived TNM classification, AJCC 7th edition (2010-2015) was limited to "T4d", SEER Combined T (2016-2017) was limited to "c2D" and "p4D", and EOD 8th edition (2018+) was limited to "T4D"; and (3) Histology type (ICD-O-3) ranged from 8500 to 8549. Year of diagnosis was limited to 2010 to 2021 because SEER started collecting breast cancer molecular subtype in 2010. To limit our data to adult females, age at diagnosis was limited to patents 20 years and older and sex was limited to "Female".

## Variables

For each case, the following information from SEER was obtained: age at diagnosis (20-49, 50-69, and \$\ge\$70), race (white, black, other), marital status (unmarried, married), N stage (N0, N1, N2, N3), M stage (M0, M1), grade (I/II, III/IV), radiation (no/unknown, yes), chemotherapy (no/unknown, yes), surgery (no/unknown, yes), survival months, cause-specific death classification (Alive or dead of other cause, Dead (attributable to this cancer dx), Dead (missing/unknown COD)), and other cause of death classification (Alive or dead due to cancer, Dead (attributable to causes other than this cancer dx), Dead (missing/unknown COD)). In addition, cases were excluded from the study if they had incomplete information on any of these characteristics. Patients who survived less than one month were also excluded. To limit our analysis to breast cancer-specific mortality, patients with unknown cause of death or patients with cause of death to another casue were excluded.

## Statistical Analysis

# Results

## Clinicopathological characteristics

## Univariate Analysis

## Multivariate Analysis

# Discussion